---
title: "Untitled"
author: "Bruno Carlin"
date: "2019-07-13"
output: html_document
---

# Basic R operations

## Vector math

No recycling

```{r}
vector1 <- c(2,3,4,6)
vector2 <- c(2,5,8,9)

vector1 + vector2
vector1 - vector2
vector1 * vector2
vector1 / vector2
vector1 ^ vector2
```

WIth recycling

```{r warning=FALSE}
vector1 <- c(2,3,4,6)
vector2 <- c(2,5,8,9,10)

vector1 + vector2
vector1 - vector2
vector1 * vector2
vector1 / vector2
vector1 ^ vector2

```


## Vector comparisson 

```{r}
c(3, 4-1,1+1+1) == 3
c(3, 4-1,1+1+1) == c(1,2,3)
1:3 != 3:1
```

## Exponential and log

```{r}
exp(1:10)
exp(1:5) < 100
```

## String comparisson

```{r}
c("A","B","C") # Boring to write
letters[1:3]
LETTERS[1:3]
letters[1:3] == LETTERS[1:3]
```

## Special numbers

```{r}
infinity <- c(Inf,-Inf)
infinity - 1
infinity + 1
infinity ^ 2
infinity + NA
infinity * NA
infinity + NaN
infinity/0
0/0
```


##  Logic Math

```{r}
z <- 1:10 >= 5
w <- 1:10 <= 3
y <- 1:10 <= 7
!z # negation
z & w # and
z || w # Strange place here.
z | w # or
z + w # conversion
(z & y) | w # combination of vectors
```

## Variable analysis

```{r}
vector1
w
class(vector1)
class(w)
str(w)
is.vector(vector1)
```

## Factors easier to read in my opinio

```{r}
library(tidyverse)
```

```{r}
sex <- c("M","F","F","aaa",NA) %>% as.factor()
sex %>% as.integer()
sex %>% levels()
```

## Random numbers

```{r}
set.seed(1)
runif(10)
runif(10)
```

```{r}
objects()
```

## Other more verbose syntax can be helpfull to readers

```{r}
remove(list = objects()) # instead of
rm(list = ls())
```

## Slicing

```{r}
character(5)
fruits <- fruit
fruits %>% class()
fruits %>% str()
fruits %>% glimpse()
fruits %>% summary()
fruits[1]
fruits[-1] %>% summary()
fruits[c(1,5,10)]
fruits[c(-1:-70)]
```

## Finding Values

```{r}
x <- -5:10
x > 5
class(x > 5)
length(x > 5)
which(x > 5)
x[(which(x > 5))]
x[x > 5]
x %>% which.max()
x %>% which.min()
```

## Repetition

```{r}
x <- 1:5
x %>% rep(each = 3)
x %>% rep(times = 1:5)
x %>% rep(times = 3)
x %>% rep(len = 4) 
```

# Matrixes

```{r}
matrix1 <- matrix(data = 1:12,
                  nrow = 4,
                  dimnames = list(c("L1","L2","L3","L4"),
                                  c("C1","C2","C3")))
matrix1
matrix1 %>% ncol()
matrix1 %>% nrow()
matrix1 %>% dim()
matrix1 %>% length()
matrix1 %>% colnames()
matrix1 %>% row.names()
colnames(matrix1) <- c("A","B","C")
matrix1 %>% colnames()
```

## Slicing


```{r}
matrix1[1,1]
matrix1[1,]
matrix1[,1]
matrix1[,"A"]
matrix1[1:2,c("C")]
```

```{r}
matrix2 <- matrix(data = seq(from = 2,to = 24,by = 2),
                  nrow = 4,
                  dimnames = list(c("L5","L6","L7","L8"),
                                  c("C5","C6","C7")))
colnames(matrix1) <- c("C1","C2","C3")
matrix1
matrix2
rbind(matrix1,matrix2)
cbind(matrix1,matrix2)

matrix1 %>% cbind(matrix2)

matrix1 + matrix2
matrix1 - matrix2
matrix1 * matrix2
matrix1 / matrix2
# matrix1 %*% matrix2 if they were compatible
t(matrix2)
matrix1 %*% t(matrix2)
diag(1,4)
```

# Data Frame

```{r}
set.seed(8)
df1 <- data.frame(x = letters [1:5],
                  y = rnorm(5),
                  z = runif(5)> 0.5,stringsAsFactors = F)
# OR

df1 <- tibble(x = letters [1:5],
                  y = rnorm(5),
                  z = runif(5)> 0.5)

df1
glimpse(df1)
class(df1)
```

## Other cool packages

```{r}
library(skimr)
df1 %>% skimr::skim()
```

```{r}
library(DataExplorer)
df1 %>% DataExplorer::introduce()
df1 %>% DataExplorer::plot_intro()
```

# More advanced commands

## If else

```{r}
if (FALSE) {
  message("this is not fake")
} else {
  print("this is fake")
  message("this is fake indeed")
}
```

### Functional if else

```{r}
0 == FALSE # R conversion
1 == TRUE  # R conversion 
set.seed(1)
coin_toss <- rbinom(10,1,0.5)
coin_toss
ifelse(test = coin_toss,
       yes = "heads",
       no = "tails")
# or
if_else(condition = coin_toss %>% as.logical(),
        true = "heads",
        false = "tails")
```

## Repetiton  structures

```{r}
action <- c("Learn R",
            "Drink Coffe",
            "Watch movies",
            "Pass in predictive models")

sample(action,size = 3,replace = TRUE)
```

### Repeat

```{r}
repeat{
  message1 <- sample(action,1)
  message(message1)
  if (message1 == "Pass in predictive models") {
    break
  }
}
```

### While

```{r}
message2 <- character(1)
while (message2 != "Pass in predictive models") {
  message2 <- sample(action,1)
  message(message2)
}
```

### For

```{r}
for (i in action) {
  message(i)
}
# or
for (i in 1:4) {
  message(i)
  message(action[[i]])
}
```

### My favourite map or apply
```{r}
action %>% str_c(" ") %>% message()
```

```{r}
fibonacci <- function(n) {
  fibo <-numeric(n)
  fibo[1] <- 1
  fibo[2] <- 1
  if (n == 0) {
    return(0)
  } else if (n == 1) {
    return(fibo[1])
  } else if (n == 2) {
    return(c(fibo[1], fibo[2]))
  } else {
    for (i in 3:n) {
      fibo[i] <- fibo[i - 2] +  fibo[i - 1]
    }
    return(fibo)
  }
}

```

```{r}
vetor <- 1:10

```

