---
title: "class_1_predictive_models"
output: html_document
---

# Predictive Model 2019-07-12

## Basic Math

```{r}
2+2
7*8
exp(1)
2^2
```

## Getting Help and commeting code

```{r}
#?log
```

## Assignement

```{r}
x <- 2 + 2
y <- x^2 + 4
```

## Clearing Workspace

```{r}
ls()
rm(list = ls())
```

## Vectors

```{r}
z <- c(1,3,4,5)
z + 1
```

```{r}
z[2:3]
```

```{r}
x <- 1:10
y <- 2*x +10
```

## Plotting

```{r}
plot(x,y)
```

### The seq operator

```{r}
seq(1,10,1)
```

```{r}
conta <- 1:100
for (i in conta) {
  conta[i] <- conta[i] +1
}
conta
```

# Inference vs prediction

## Example of a linear model

y = -80 + 9.7 * X

```{r}
X <- 10
y_hat <- -80 + 9.7* X
y_hat
```

## Error functions

```{r}
y <- 15
error <- y - y_hat
error

MASE <- error^2
MASE

MAPE <-  abs(error)
MAPE
```
When doing models we want to minimize our EQM function

$$ 
EQM = bias(x)^2 + var(x)
$$

$$
\hat\psi(x) = Best Calculated EQM
$$

## Distributions

### Setting Seed

```{r}
set.seed(123)
```

### Uniform Distribution
```{r}
x <- runif(n = 30,min = 8,max = 18)
x
```

### Functional programming

```{r}
library(magrittr)
x %>% sort()
```

```{r}
y <- 45 * tanh(x/1.9 - 7) + 57 + rnorm(30,mean = 0,sd = 4)
```

```{r}
library(ggplot2)
library(dplyr)
```

```{r}
df <- tibble(x,y)
```

```{r}
curve(45 * tanh(x/1.9 - 7) + 57,from = 8,to = 18,lwd = 2)
points(x,y,col = "blue",pch = 19)
```

```{r}
tahn1  = function(x){
  45 * tanh(x/1.9 - 7) + 57
}
```


```{r}
ggplot(df) +
  aes(x,y) +
  geom_point() +
  stat_function(fun = tahn1,size =1,color = "blue") +
  geom_line(data = data.frame(spline(x,y,2)),color = "red") +
  xlim(8,18)
```

## Simple Model

### Fit and plot

```{r}
set.seed(123)

x <- runif(n = 30,min = 8,max = 18) %>% sort()
y <- 45 * tanh(x/1.9 - 7) + 57 + rnorm(30,mean = 0,sd = 4)

y_hat <- numeric()

curve(45 * tanh(x/1.9 - 7) + 57,from = 8,to = 18,lwd = 2)

for(i in 1:500){
  x_sim <- runif(n = 30,min = 8,max = 18) %>% sort()
  y_sim <- tahn1(x_sim)+ rnorm(30,mean = 0,sd = 4)
  
  y_hat[i] <- mean(y_sim)
  abline(h = y_hat[i],col = "grey", lwd = 2)
  abline(h = y_hat %>% mean(),col = "red",lwd = 2)
  
}
```


### Bias and variance against fixed point

```{r}
(bias_11 <- y_hat %>% mean() - tahn1(11))
(bias_15 <- y_hat %>% mean() - tahn1(15))
(variance_15 <- variance_11  <- y_hat %>% var())
(EQM_11 <- bias_11 ^ 2 + variance_11)
(EQM_15 <- bias_15 ^ 2 + variance_15)
```


## A more flexible model, spline 2

### Fit and plot

```{r}
set.seed(123)

x <- runif(n = 30,min = 8,max = 18) %>% sort()
y <- 45 * tanh(x/1.9 - 7) + 57 + rnorm(30,mean = 0,sd = 4)

y_hat15 <- y_hat11 <- y_hat <- numeric()

curve(45 * tanh(x/1.9 - 7) + 57,from = 8,to = 18,lwd = 2)

for(i in 1:500){
  x_sim <- runif(n = 30,min = 8,max = 18) %>% sort()
  y_sim <- tahn1(x_sim)+ rnorm(30,mean = 0,sd = 4)
  
  fit <- smooth.spline(x_sim,y_sim,df = 2)
  lines(predict(fit,seq(8,18,0.5)),col = "grey",lwd = 2)
  
  y_hat11[i] <- predict(fit,x =11)$y
  y_hat15[i] <- predict(fit,x =15)$y
}
  points(11, mean(y_hat11), pch = 18, col = "red")
  points(15, mean(y_hat15), pch = 18, col = "red")

```

### Bias and Variance

```{r}
(bias_11 <- (y_hat11 %>% mean() - tahn1(11)) %>% `^`(2))
(bias_15 <- (y_hat15 %>% mean() - tahn1(15)) %>% `^`(2))
(variance_11  <- y_hat11 %>% var())
(variance_15 <- y_hat15 %>% var())
(EQM_11 <- bias_11 + variance_11)
(EQM_15 <- bias_15 + variance_15)
```

## A even more flexible model spline = 7

### Fit and plot

```{r}
set.seed(123)

x <- runif(n = 30,min = 8,max = 18) %>% sort()
y <- 45 * tanh(x/1.9 - 7) + 57 + rnorm(30,mean = 0,sd = 4)

y_hat15 <- y_hat11 <- y_hat <- numeric()

curve(45 * tanh(x/1.9 - 7) + 57,from = 8,to = 18,lwd = 2,col = "black")

for(i in 1:500){
  x_sim <- runif(n = 30,min = 8,max = 18) %>% sort()
  y_sim <- tahn1(x_sim)+ rnorm(30,mean = 0,sd = 4)
  
  fit <- smooth.spline(x_sim,y_sim,df = 7)
  lines(predict(fit,seq(8,18,0.5)),col = "grey",lwd = 2)
  
  y_hat11[i] <- predict(fit,x =11)$y
  y_hat15[i] <- predict(fit,x =15)$y
}
  points(11, mean(y_hat11), pch = 18, col = "red")
  points(15, mean(y_hat15), pch = 18, col = "red")
```

### Bias and Variance

```{r}
(bias_11 <- (y_hat11 %>% mean() - tahn1(11)) %>% `^`(2))
(bias_15 <- (y_hat15 %>% mean() - tahn1(15)) %>% `^`(2))
(variance_11  <- y_hat11 %>% var())
(variance_15 <- y_hat15 %>% var())
(EQM_11 <- bias_11 + variance_11)
(EQM_15 <- bias_15 + variance_15)
```


## An even too flexible model spline = 20

### Fit and plot

```{r}
set.seed(123)

x <- runif(n = 30,min = 8,max = 18) %>% sort()
y <- 45 * tanh(x/1.9 - 7) + 57 + rnorm(30,mean = 0,sd = 4)

y_hat15 <- y_hat11 <- y_hat <- numeric()

curve(45 * tanh(x/1.9 - 7) + 57,from = 8,to = 18,lwd = 2,col = "black")

for(i in 1:500){
  x_sim <- runif(n = 30,min = 8,max = 18) %>% sort()
  y_sim <- tahn1(x_sim)+ rnorm(30,mean = 0,sd = 4)
  
  fit <- smooth.spline(x_sim,y_sim,df = 20)
  lines(predict(fit,seq(8,18,0.5)),col = "grey",lwd = 2)
  
  y_hat11[i] <- predict(fit,x =11)$y
  y_hat15[i] <- predict(fit,x =15)$y
}
  points(11, mean(y_hat11), pch = 18, col = "red")
  points(15, mean(y_hat15), pch = 18, col = "red")
```

### Bias and Variance

```{r}
(bias_11 <- (y_hat11 %>% mean() - tahn1(11)) %>% `^`(2))
(bias_15 <- (y_hat15 %>% mean() - tahn1(15)) %>% `^`(2))
(variance_11  <- y_hat11 %>% var())
(variance_15 <- y_hat15 %>% var())
(EQM_11 <- bias_11 + variance_11)
(EQM_15 <- bias_15 + variance_15)
```

## An even way too flexible model spline = 29

### Fit and plot

```{r}
set.seed(123)

x <- runif(n = 30,min = 8,max = 18) %>% sort()
y <- 45 * tanh(x/1.9 - 7) + 57 + rnorm(30,mean = 0,sd = 4)

y_hat15 <- y_hat11 <- y_hat <- numeric()

curve((45 * tanh(x/1.9 - 7) + 57),from = 8,to = 18,lwd = 2,col = "black")

for(i in 1:500){
  x_sim <- runif(n = 30,min = 8,max = 18) %>% sort()
  y_sim <- tahn1(x_sim)+ rnorm(30,mean = 0,sd = 4)
  
  fit <- smooth.spline(x_sim,y_sim,df = 29)
  lines(predict(fit,seq(8,18,0.5)),col = "grey",lwd = 2)
  
  y_hat11[i] <- predict(fit,x =11)$y
  y_hat15[i] <- predict(fit,x =15)$y
}
  points(11, mean(y_hat11), pch = 18, col = "red")
  points(15, mean(y_hat15), pch = 18, col = "red")
```

### Bias and Variance

```{r}
(bias_11 <- (y_hat11 %>% mean() - tahn1(11)) %>% `^`(2))
(bias_15 <- (y_hat15 %>% mean() - tahn1(15)) %>% `^`(2))
(variance_11  <- y_hat11 %>% var())
(variance_15 <- y_hat15 %>% var())
(EQM_11 <- bias_11 + variance_11)
(EQM_15 <- bias_15 + variance_15)
```
